Cloud correction and filtering of operational microwave humidity radiances with case specific uncertainty estimation

### 2020-09-24 Article/ici_submm_cloudcorr.tex
some minor changes

This study presents a cloud filtering and correction method for microwave humidity radiances. The method is based on the postulation that sub-millimeter(sub-mm) frequencies can provide a basis for cloud filtering/correction of data measured around 183\,GHz. A quantile regression(QRNN) based approach is formulated to estimate the posterior distributions of noise free clear sky (NFCS) radiances. QRNNs are trained to correct the cloud impact for each 183\,GHz channel using measurements from the target channel and available sub-mm channels. The approach is demonstrated by application to Ice Cloud Imager(ICI) and MicroWave Imager (MWI). As a special case applicability to smaller satellite missions like Arctic Weather Satellite (AWS), with fewer sub-mm channels, is also evaluated. The performance of QRNN based cloud filtering is also investigated. The results are compared against the NFCS simulations.

@@

The results show that it can be highly recommended to use all available sub-mm frequencies to estimate the NFCS values of radiances/brightness temperatures. The strategy works well even when only one of the higher frequencies is available. This study uses simulated data, but we can expect to obtain similar performance when data fron the new sensors will be available.

@@

\subsubsection{Analysis of cloud filtering}
%
\label{sec:cloud_filtering}
%f 
\begin{figure*}[t ]
	\includegraphics[width=\textwidth]{Figures/error_distribution_QRNN-single.pdf} 
	\caption{Error distributions for deviations of filtered and predicted data to clear-sky simulations. The results are from QRNN-single and channels I1V, I2V and I3V. Noise for each channel is also plotted for reference. The label ``Uncorrected'' represents the measurement data and ``Filtered(5\,K)'' represents the dataset when no correction is made but cases with cloud correction greater than 5\,K are rejected. ``Predicted(All)'' denotes the entire prediction dataset, while ``Predicted(5\,K)'' refers to the predicted dataset but where cases with cloud correction greater than 5\,K are excluded.}
	\label{fig:error_distributions}	
\end{figure*}
\begin{table*}[t]
	\caption{ Error statistics for deviations of filtered and predicted data to NFCS simulations. 
		The values for Bias, mean absolute error(MAE), standard deviation(STD), and measure of skewness(Skewness) are shown. Results are from both QRNN-single and QRNN-all for channels I1V, I2V and I3V. The number in parentheses is the percentage of cases removed by filtering. The label ``Filtered(5\,K)'' represents the dataset when no correction is made but cases with cloud correction greater than 5\,K are rejected, ``Predicted(All)'' refers to the complete predicted dataset, while ``Predicted (5\,K)'' denotes the prediction dataset but where cases with cloud correction greater than 5\,K are excluded.}
	\label{tab:error_statistics_ici}
	%	\tabcolsep=0.11cm
	\begin{tabular}{llrr|rrr|rrr}
		\tophline
		&&\multicolumn{2}{c|}{Simulations}& \multicolumn{3}{c|}{QRNN-single} & \multicolumn{3}{c}{QRNN-all}\\
		\cline{3-10}
		%		\hline
		&&   Clear-sky &   All-sky &  Filtered & Predicted & Predicted &   Filtered & Predicted & Predicted \\
		&&&&							(5\,K) &(All)& (5\,K) & (5\,K)&(All)& (5\,K)\\
		\middlehline
		%		\multicolumn{7}{c}{Channel - I1V}\\

@@

To implement the cloud filtering scheme, we chose 5\,K cloud correction as the filtering threshold. All cases with correction greater than 5\,K were rejected. In order to analyse the performance of the filtering threshold, the deviations of the filtered measurements to the NFCS simulations were calculated. Fig.~\ref{fig:error_distributions} shows the distribution of the deviations for filtered(red) and original datasets(green). Cloud contamination causes a reduction in the brightness temperatures, and introduces a large negative bias in the error distributions. Without any cloud impact, the error distributions must follow a Gaussian distribution with a narrow spread. When filtering is activated, the resulting distributions have a low but asymmetric spread. The right side of the distributions follows the measurement noise, as expected. The left side has a significantly smaller negative skewness than the measurements. This indicates that 5\,K filter can successfully remove cases associated with high cloud impact. However at the same time, the asymmetric error distribution indicates that cases with low cloud impact are erroneously classified as clear-sky.

For a quantitative assessment of the performance achieved with cloud filtering, we calculated the average error bias, MAE (mean absolute error) and standard deviation of the errors. The asymmetry of error distributions around their mean is also calculated through the measure of skewness. Fisher-Pearson coefficient of skewness is used to calculate the measure of skewness: 
\begin{equation}
g_1 = \frac{m_3}{m_2^{3/2}}, 
\end{equation}
where $m_2$ and $m_3$ are biased sample's second and third central moments. The error metrics corresponding to the error distributions described in Fig.~\ref{fig:error_distributions} are provided in Table~\ref{tab:error_statistics_ici} under label "Filtered(5\,K)". The average bias in I1V measurements is -1.88\,K, and rejecting around 6\% of the cases as cloudy reduces the bias to -0.32\,K. The standard deviation also reduces to 1.07\,K from 8.84\,K. Similarly for I2V, the average error bias in all sky measurements is -1.04\,K, that reduces to -0.24\,K after filtering. The corresponding improvement in MAE is almost 65\%. Only 3.7\% of the total cases are rejected to achieve this accuracy. For I3V, filtering classifies 2.3\% of the total cases as cloudy and reduces the bias to 0.18\,K from -0.63\,K and the standard deviation from 5.95\,K to 0.91\,K. The results from QRNN-all are also shown in the same table. Using all 183\,GHz channels concurrently has no notable impact on the cloud filtering. Almost similar statistics are obtained from QRNN-single and QRNN-all. Overall for all three channels, the notable decrease in the errors indicates that the cloud filter is effective in removing most of the cloudy cases, however few cases with low cloud impact still pass through the filter and deteriorate the performance. Thus pure-filtering approach can prove as a reliable procedure for removing cases with high cloud impact, but additional measures are needed to remove cases with low cloud impact.

>>>>

This study presents a cloud filtering and correction method for microwave humidity radiances. The method is based on the postulation that sub-millimeter(sub-mm) frequencies can provide a basis for cloud filtering/correction of data measured around 183\,GHz. A quantile regression(QRNN) based approach is formulated to estimate the posterior distributions of noise free clear sky (NFCS) radiances. QRNNs are trained to correct the cloud impact for each 183\,GHz channel using measurements from the target channel and available sub-mm channels. The approach is demonstrated by application to Ice Cloud Imager(ICI) and MicroWave Imager (MWI). As a special case, applicability to smaller satellite missions like Arctic Weather Satellite (AWS), with fewer sub-mm channels, is also evaluated. The performance of QRNN based cloud filtering is also investigated. The results are compared against the NFCS simulations.

@@

The results show that it can be highly recommended to use all available sub-mm frequencies to estimate the NFCS values of radiances/brightness temperatures. The strategy works well even when only one of the higher frequencies is available. This study uses simulated data, but we can expect to obtain similar performance when data from the new sensors will be available.

@@

\subsubsection{Analysis of cloud filtering}
%
\label{sec:cloud_filtering}
%f 
\begin{figure*}[t ]
	\includegraphics[width=\textwidth]{Figures/error_distribution_QRNN-single.pdf} 
	\caption{Error distributions for deviations of filtered and predicted data to NFCS simulations. The results are from QRNN-single and channels I1V, I2V and I3V. Noise for each channel is also plotted for reference. The label ``Uncorrected'' represents the measurement data and ``Filtered(5\,K)'' represents the dataset when no correction is made but cases with cloud correction greater than 5\,K are rejected. ``Predicted(All)'' denotes the entire prediction dataset, while ``Predicted(5\,K)'' refers to the predicted dataset but where cases with cloud correction greater than 5\,K are excluded.}
	\label{fig:error_distributions}	
\end{figure*}
\begin{table*}[t]
	\caption{ Error statistics for deviations of filtered and predicted data to NFCS simulations. 
		The values for Bias, mean absolute error(MAE), standard deviation(STD), and measure of skewness(Skewness) are shown. Results are from both QRNN-single and QRNN-all and for channels I1V, I2V and I3V. The number in parentheses is the percentage of cases removed by filtering. The label ``Filtered(5\,K)'' represents the dataset when no correction is made but cases with cloud correction greater than 5\,K are rejected, ``Predicted(All)'' refers to the complete predicted dataset, while ``Predicted (5\,K)'' denotes the prediction dataset but where cases with cloud correction greater than 5\,K are excluded.}
	\label{tab:error_statistics_ici}
	%	\tabcolsep=0.11cm
	\begin{tabular}{llrr|rrr|rrr}
		\tophline
		&&\multicolumn{2}{c|}{Simulations}& \multicolumn{3}{c|}{QRNN-single} & \multicolumn{3}{c}{QRNN-all}\\
		\cline{3-10}
		%		\hline
		&&   Clear-sky &   All-sky &  Filtered & Predicted & Predicted &   Filtered & Predicted & Predicted \\
		&&&&							(5\,K) &(All)& (5\,K) & (5\,K)&(All)& (5\,K)\\
		\middlehline
		%		\multicolumn{7}{c}{Channel - I1V}\\

@@

To implement the cloud filtering scheme, we chose 5\,K cloud correction as the filtering threshold. All cases with correction greater than 5\,K were rejected. In order to analyse the performance of the filtering threshold, the deviations of the filtered measurements to the NFCS simulations were calculated. Fig.~\ref{fig:error_distributions} shows the distribution of the deviations for filtered(red) and original datasets(green). Cloud contamination causes a reduction in the brightness temperatures, and introduces a large negative bias in the error distributions. Without any cloud impact, the error distributions must follow a Gaussian distribution with a narrow spread. When filtering is activated, the resulting distributions have a low but asymmetric spread. The right side of the distributions follows the measurement noise, as expected, but the left side has a significantly smaller negative skewness than the measurements. This indicates that 5\,K filter can successfully remove cases associated with high cloud impact. However at the same time, the asymmetric error distribution is a result of cases erroneously classified as clear-sky.

For a quantitative assessment of the performance achieved with cloud filtering, we calculated the average error bias, MAE (mean absolute error) and standard deviation of the deviations. The asymmetry of error distributions around their mean is also calculated through the measure of skewness. Fisher-Pearson coefficient of skewness is used to calculate the measure of skewness: 
\begin{equation}
g_1 = \frac{m_3}{m_2^{3/2}}, 
\end{equation}
where $m_2$ and $m_3$ are biased sample's second and third central moments. The error metrics corresponding to the error distributions described in Fig.~\ref{fig:error_distributions} are provided in Table~\ref{tab:error_statistics_ici} under label "Filtered(5\,K)". The average bias in I1V measurements is -1.88\,K, and rejecting around 6\% of the cases as cloudy, reduces the bias to -0.32\,K. The standard deviation also reduces to 1.07\,K from 8.84\,K. Similarly for I2V, the average error bias in all sky measurements is -1.04\,K, that reduces to -0.24\,K after filtering. The corresponding improvement in MAE is almost 65\%. Only 3.7\% of the total cases are rejected to achieve this accuracy. For I3V, filtering classifies 2.3\% of the total cases as cloudy and reduces the bias to 0.18\,K from -0.63\,K and the standard deviation from 5.95\,K to 0.91\,K. The results from QRNN-all are also shown in the same table. Using all 183\,GHz channels concurrently has no notable impact on the cloud filtering. Almost similar statistics are obtained from both models.

Overall for all three channels, the notable decrease in the errors indicates that the cloud filter is effective in removing most of the cloudy cases, however few cases with low cloud impact still pass through the filter and deteriorate the performance. Thus pure-filtering approach can prove as a reliable procedure for removing cases with high cloud impact, but additional measures are needed to remove cases with low cloud impact.
